```{r}
library(tidyverse)
```

```{r}
matchsets <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/main/data/ncaa_womens_volleyball_matchstats_2022.csv")
```

```{r}
teamquality <- matchsets |> 
  select(team, defensive_r_err, defensive_b_err, tb, hit_pct, total_attacks, team_score, opponent_score)
```

```{r}
teamtotals <- teamquality |> 
  group_by(team) |> 
  summarise(
    DefRErrAvg = mean(defensive_r_err), 
    DefBErrAvg = mean(defensive_b_err), 
    TBAvg = mean(tb),
    HitPctAvg = mean(hit_pct),
    TotalAttackAvg = mean(total_attacks),
    TeamScoreAvg = mean(team_score),
    OppScoreAvg = mean(opponent_score)
    ) 
```

```{r}
teamzscorevball <- teamtotals |>
  mutate(
    TBZscore = as.numeric(scale(TBAvg, center = TRUE, scale = TRUE)),
    HitPctZscore = as.numeric(scale(HitPctAvg, center = TRUE, scale = TRUE)),
    TotalAttackZscore = as.numeric(scale(TotalAttackAvg, center = TRUE, scale = TRUE)),
    TeamScoreZscore = as.numeric(scale(TeamScoreAvg, center = TRUE, scale = TRUE)),
    OppScoreZscore = as.numeric(scale(OppScoreAvg, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = TBZscore + HitPctZscore + TotalAttackZscore + TeamScoreZscore + OppScoreZscore
    ) |> ungroup()
```

```{r}
head(teamzscorevball)
```

```{r}
teamzscorevball |> arrange(desc(TotalZscore))
```
Notes: 
Maryland is closer to a 0 z-score, which means that Maryland is much closer to the mean of the data score, with a slight over performance in team quality. 

I what is interesting in this data, is Wisconsin's effectiveness as a team. There are here at #1, outperforming other teams. This data is from 2022, and this year Wisconsin is #1 so far. Although this is not a predictive tool like residuals, I find this an interesting finding that I wonder if these metrics are contributing to their success this year. 

Write-up:
The two effective college volleyball teams, Wisconsin and Towson, are out dominating there respective conferences (NCAA and CAA). Taking a look at the teams total blocks, hit percentages, total attacks and team and opponent scores, we can determine the team's quality. 

With the mentioned merits averaged together, Wisconsin and Towson take the #1 and #2 spots in our data set. In comparison with other teams, these two have a higher block to attack ratio, meaning together Wisconsin and Town will average around 120 for attacks and 10.5 for blocks. Whereas in lower positions this ratio is much lower. For example the Manahatten Jaspers will have a 1.9 block average and 85.9 attack average. See the difference? 

(I think here I would still somehow like to explore how to talk about Wisconsins potential rise to the top next season - 2023).  

```{r}
 marylandzscore <- teamzscorevball |> 
     group_by(team) |>
  filter(TotalZscore > 5) |>
  arrange(desc(TotalZscore))
```

I think a good story here would be the difference between Towson's success and Maryland's success within these merits. Towson is within the top 5 whereas Maryland is bordering the mean. We've discussed in class the geography of the team explaining an undoubted success for the team. But here we have two teams about an hour away from one another, I would like to investigate this more, and its potential differences from season to season. 

#scraping data 
```{r}
 vbdata <- read_csv("https://raw.githubusercontent.com/Sports-Roster-Data/womens-volleyball/main/vb_rosters_2023_24.csv")
```

```{r}
rosters_with_towns <- rosters |>
  mutate(HomeTownCopy = Hometown)
  separate(Hometown, c('town', 'state'),sep=',')
```


##Dumbbell and Lollipop Charts

```{r}
library(tidyverse)
library(ggalt)
```
```{r}
logs <- read_csv("https://dwillis.github.io/sports-data-files/ncaa_womens_volleyball_matchstats_2022.csv")
```
Looking at the difference between a team’s errors versus their opponents’ errors. We’re going to add up all team errors and opponent errors for a team in a season and take a look at where they come out. 
```{r}
big10 <- c("Nebraska Cornhuskers", "Iowa Hawkeyes", "Minnesota Golden Gophers", "Illinois Fighting Illini", "Northwestern Wildcats", "Wisconsin Badgers", "Indiana Hoosiers", "Purdue Boilermakers", "Ohio St. Buckeyes", "Michigan Wolverines", "Michigan St. Spartans", "Penn St. Nittany Lions", "Rutgers Scarlet Knights", "Maryland Terrapins, Terps")

errors <- logs |>
  filter(team %in% big10) |> 
  group_by(team) |> 
  summarise(
    total_errors = sum(errors), 
    opp_errors = sum(defensive_errors))
```

Let's create the plot! xend= end of the dumbbell. Lets distinguish opponent errors from team errors with color, makes the dots larger, and lighten the background. 
```{r}
ggplot() + 
  geom_dumbbell(
    data=errors, 
    aes(y=team, x=total_errors, xend=opp_errors),
    size = 1,
    colour = "grey",
    colour_x = "green",
    colour_xend = "red") + 
  theme_minimal()
  
```


```{r}
ggplot() + 
  geom_dumbbell(
    data=errors, 
    aes(y=reorder(team, opp_errors), x=total_errors, xend=opp_errors),
    size = 1,
    colour = "grey",
    colour_x = "green",
    colour_xend = "red") + 
  theme_minimal()
```
Looking at the graph this way we can see that Maryland's opponents made the most errors this season. You can also see that the best teams: Michigan, Wisconsin and Minnesota - dont make too many errors.


Lollipop Charts
looking specifically at opponent errors for teams. 
```{r}
ggplot() + 
  geom_lollipop(
    data=errors, 
    aes(y=team, x=opp_errors), 
    horizontal = TRUE
    ) + theme_minimal() + 
  labs(title = "Maryland, Rutgers force more errors", y="team")
```

```{r}
md <- errors |> filter(team == "Maryland Terrapins, Terps")
```
```{r}
ggplot() + 
  geom_lollipop(
    data=errors, 
    aes(y=reorder(team, opp_errors), x=opp_errors), 
    horizontal = TRUE
    ) + 
  geom_lollipop(
    data=md,
    aes(y=team, x=opp_errors),
    horizontal = TRUE,
    color = "red"
  ) + 
  theme_minimal() + 
  labs(title = "Maryland forced the most errors among Big Ten teams", y="team")
```

