Should I change to: 
Individual players and when are successful players at their most successful/relied upon? 
Or can just expand upon my initial question, Do successful (winning) Big Ten teams rely on their first-year players more than others? When are freshman players relied upon? 

College career (Freshman -> Senior) playtime and success? 

##Data Backgrounding 
Do successful (winning) Big Ten teams rely on their first-year players more than others? 

```{r}
library(tidyverse)
```

Data needed: Players and class standings, Shots by each player (and will compare by their class standings), Goals by each player, Assists by each player, Position (I don't think this is necessary for my question but I think can lead to interesting results and may help me to further expand my analysis to other stories/questions)

```{r}
umd_players_2022 <- read_csv ("https://github.com/gryan517/-jour479x_fall_2023/blob/main/data/umd_2022_player_stats.csv")
```

I'm having trouble figuring out the "Rows: 0", but this is what the step would look like once I figure this out.

```{r}
head(umd_players_2022)
```

```{r}
umd_gameresults_2023 <- read_csv ("https://github.com/gryan517/-jour479x_fall_2023/blob/7cb7b8612eb310502f086ea26d3032c20871c854/data/umd_2022_gameresults.csv")
```
I feel that there is a more simplified way of inputting the data. So before doing this for each Big Ten team, I think this is my biggest question when inputting the data. I have all the data I need here my questions for it are: Will the game results be interpreted by R in the "GOALS SCORED [ASSISTED]" column? And will the data still be plausible to use if I enter it into one csv file?

What I would do with this data is take a look at the overall amount of goals made by freshman in comparison to other class standings, as well as how many of those goals were assisted by other freshman or class standing. Additionally I would look to the winning games and who made goals in that game (but the way that Maryland has the data format, I'm assuming that R wont be able to interpret this?). 

An aside, I still think it would be interesting to include position as I think this will lead to interesting discoveries once I start more analyzing.
