
Or can just expand upon my initial question, Do successful (winning) Big Ten teams rely on their first-year players more than others? When are freshman players relied upon? 
 

##Data Backgrounding 
Do successful (winning) Big Ten teams rely on their first-year players more than others? 

```{r}
library(tidyverse)
```

Data needed: Players and class standings, Shots by each player (and will compare by their class standings), Goals by each player, Assists by each player, Position (I don't think this is necessary for my question but I think can lead to interesting results and may help me to further expand my analysis to other stories/questions)

```{r}
ncaa_players_2022 <- read_csv ("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_playerstats_2022.csv")
```

```{r}
head(ncaa_players_2022)
```
```{r}
ncaa_players_2022|>
  group_by(team, yr) |>
  summarise(
    total = n()
  ) 
```
```{r}
ncaa_players_2022|>
  group_by(team, yr) |>
  summarise(
    total = n(),
    mean_goals = mean(goals)
  )
```

Basically what I want to see is from each team, the avg amount of goals for each yr by team. I wouldn't do it for the NCAA entirely because 1. Looking at the teams separately tells more of a story (we can see if one team relies on one year more than another) 2. We can easily narrow it down to the big 10 teams. 

I think the above can be a way to do this. I wonder if I am getting somewhere with the coding below? I'm hitting a road block though.


```{r}
byyr <- ncaa_players_2022 |>
  mutate(
    goalspergame = goals/games
  ) |> 
  filter(
    yr == "Fr" |
    yr == "Sr" |
    yr == "So" | 
    yr == "Jr")
 
```

```{r}
ggplot() + geom_bar(data=byyr, aes(x=yr, weight=goalspergame))
```
This is for all of NCAA grouped together by yr. Now I just have to filter by the big ten teams. Also this doesnt seem right, or maybe I am interpreting this wrong, but the numbers seem way too big. I understand these dont represent an average for each game, but does this represent overall, across all games, each year has scored these many goals across the season on average??

```{r}
big10 <- ncaa_players_2022 |>
  mutate(
    goalspergame = goals/games
  ) |> 
  filter(
    team == "Big Ten" 
  )
```
I would like to narrow this down to big ten schools and then show maybe a stacked bar graph of each year for each school out of the big ten.I think I would need different data for this that also represents the conference. I could filter through the data, but I wasnt able to find all the bug 10 schools so I think there would be issues in this. I would also like to incorporate the (or potentially replace amount of goals) with minutes played. But ultimately once i have access to data that has the conference and minutes played, I think my graphs will look a lot different from wht it looks like now. 


##other ideas 
```{r}
pergame <- ncaa_players_2022 |>
  group_by (team) |>
  mutate(
  gspergame = gs/games,
  goalspergame = goals/games,
  pointspergame = points/games, 
  assistspergame = assists/games
  ) |>
  select(team, yr, pos, gs, goals, points, assists)
```


Im having abit of trouble with this code block, I think I need to enter na.omit(), or basically omit the NA's. Once I am abkle to do this 
```{r}
teamzscore <- pergame |>
  group_by(team)|>
  mutate(
    gszscore = as.numeric(scale(gspergame, center = TRUE, scale = TRUE)),
    goalszscore = as.numeric(scale(goalspergame, center = TRUE, scale = TRUE)),
    pointszscore = as.numeric(scale(pointspergame, center = TRUE, scale = TRUE)),
    assistszscore = as.numeric(scale(assistspergame, center = TRUE, scale = TRUE)),
    totalzscore = gszscore + goalszscore + pointszscore + assistszscore
  ) |> 
  ungroup() |> na.omit()
```

```{r}
umd_gameresults_2023 <- read_csv ("https://github.com/gryan517/-jour479x_fall_2023/blob/7cb7b8612eb310502f086ea26d3032c20871c854/data/umd_2022_gameresults.csv")
```
I feel that there is a more simplified way of inputting the data. So before doing this for each Big Ten team, I think this is my biggest question when inputting the data. I have all the data I need here my questions for it are: Will the game results be interpreted by R in the "GOALS SCORED [ASSISTED]" column? And will the data still be plausible to use if I enter it into one csv file?

What I would do with this data is take a look at the overall amount of goals made by freshman in comparison to other class standings, as well as how many of those goals were assisted by other freshman or class standing. Additionally I would look to the winning games and who made goals in that game (but the way that Maryland has the data format, I'm assuming that R wont be able to interpret this?). 

An aside, I still think it would be interesting to include position as I think this will lead to interesting discoveries once I start more analyzing.
