##Data Backgrounding 
Do successful (winning) Big Ten teams rely on their first-year players more than others? 

```{r}
library(tidyverse)
library(dplyr)
```


```{r}
ncaa_players_2023 <- read_csv ("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_playerstats_2023.csv")
```


```{r}
head(ncaa_players_2023)
```
```{r}
ncaa_players_2023|>
  group_by(team, yr) |>
  summarise(
    total = n()
  ) 
```

```{r}
ncaa_players_2023|>
  group_by(team, yr) |>
  summarise(
    total = n(),
    mean_goals = mean(goals)
  ) 
```

Basically what I want to see is from each team, the avg amount of goals for each yr by team. I wouldn't do it for the NCAA entirely because 1. Looking at the teams separately tells more of a story (we can see if one team relies on one year more than another) 2. We can easily narrow it down to the big 10 teams. 


```{r}
byyr <- ncaa_players_2023 |>
  mutate(
    goalspergame = goals/games, 
    goals <- goals %>% replace(is.na(.), 0)) |>
  filter(
    yr == "Fr" |
    yr == "Sr" |
    yr == "So" | 
    yr == "Jr") |>
  filter(
    team == "Ohio St. Buckeyes" |
    team == "Michigan St. Spartans" |
    team == "Penn St. Nittany Lions" |
    team == "Rutgers Scarlet Knights" |
    team == "Maryland Terrapins, Terps" |
    team == "Michigan St. Spartans" |
    team == "Indiana Hoosiers" |
    team == "Iowa Hawkeyes" |
    team == "Wisconsin Badgers" |
    team == "Minnesota Golden Gophers" |
    team == "Nebraska Huskers" |
    team == "Northwestern Wildcats" |
    team == "Illinois St. Redbirds" |
    team == "Purdue Boilermakers"
  )
```

```{r}
ggplot() + 
  geom_bar(
    data=byyr, 
    aes(x=team, weight=goalspergame, fill=yr)) + 
  coord_flip()
```

So now that we have a graph that represents the amount of goals each year from each team, I think this can more accurately represent the data I would like to display. Additionally, I would like to do the same with minutes played for each year for each Big Ten Team. This all would accurately align with my question. 


##assists by year
```{r}
ncaa_players_2023|>
  group_by(team, yr) |>
  mutate(assists <- assists %>% replace(is.na(.), 0)) |>
  summarise(
    total = n()
  ) 
```


```{r}
assistsbyyr <- ncaa_players_2023 |>
  mutate(
    assistspergame = assists/games, 
    goals <- goals %>% replace(is.na(.), 0)) |>
  filter(
    yr == "Fr" |
    yr == "Sr" |
    yr == "So" | 
    yr == "Jr") |>
  filter(
    team == "Ohio St. Buckeyes" |
    team == "Michigan St. Spartans" |
    team == "Penn St. Nittany Lions" |
    team == "Rutgers Scarlet Knights" |
    team == "Maryland Terrapins, Terps" |
    team == "Michigan St. Spartans" |
    team == "Indiana Hoosiers" |
    team == "Iowa Hawkeyes" |
    team == "Wisconsin Badgers" |
    team == "Minnesota Golden Gophers" |
    team == "Nebraska Huskers" |
    team == "Northwestern Wildcats" |
    team == "Illinois St. Redbirds" |
    team == "Purdue Boilermakers"
  )
```

```{r}
ggplot() + 
  geom_bar(
    data=byyr, 
    aes(x=team, weight=goalspergame, fill=yr)) + 
  coord_flip()
```
```{r}
assistsbyyr |>
  group_by(team, yr) |>
  summarise(
    sum = (assists)
  ) |>
  filter(
    yr == "Fr")
```

##joining

