##Data Backgrounding 
Do successful (winning) Big Ten/top 10 teams rely on their first-year players more than others? 

```{r}
library(tidyverse)
library(dplyr)
library(scales)
```


```{r}
ncaa_players_2023 <- read_csv ("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_playerstats_2023.csv")
```

```{r}
head(ncaa_players_2023)
```
```{r}
ncaa_players_2023|>
  group_by(team, yr) |>
  summarise(
    total = n()
  ) 
```
#The general overview

Here I am creating a graph to demonstrate the relationship between each year for each of the top 10 teams.The graph displays the amount of goals each team makes per game.

```{r}
player_goals_2023 <- ncaa_players_2023|>
  group_by(team, yr) |>
  filter(!is.na(goals))|>
   mutate(
    goalspergame = goals/games)
```

Basically what I want to see is from each team, the amount of goals for each game by team and yr. I wouldn't do it for the NCAA entirely because 1. Looking at the teams separately tells more of a story (we can see if one team relies on one year more than another) 2. We can easily narrow it down to the big 10 teams or the top 10 teams.


##Narrowing it down
```{r}
team_gpg_2023 <- player_goals_2023 |>
  group_by(team)|> 
  summarise(
    total_goals = sum(goals), 
    total_games = max(games), 
    team_gpg = total_goals/total_games) |>
  arrange(desc(team_gpg))
```

##Highest scoring offensive 
```{r}
umd_gpg_2023 <- team_gpg_2023 |> filter(team == "Maryland Terrapins, Terps")
```

```{r}
ggplot() + geom_point(data=team_gpg_2023, aes(x=total_games, y=total_goals)) +
  geom_point(data=umd_gpg_2023, aes(x=total_games, y=total_goals), color="red") +
    labs(
    title="UMD scores average Goals", 
    x="Total Goals", 
    y="Total Goals")
```

##Big Ten

```{r}
big_ten_gpg_2023 <- team_gpg_2023 |>
  group_by (team) |>
  filter(
    team == "Ohio St. Buckeyes" |
    team == "Michigan St. Spartans" |
    team == "Penn St. Nittany Lions" |
    team == "Rutgers Scarlet Knights" |
    team == "Maryland Terrapins, Terps" |
    team == "Michigan St. Spartans" |
    team == "Indiana Hoosiers" |
    team == "Iowa Hawkeyes" |
    team == "Wisconsin Badgers" |
    team == "Minnesota Golden Gophers" |
    team == "Nebraska Huskers" |
    team == "Northwestern Wildcats" |
    team == "Illinois St. Redbirds" |
    team == "Purdue Boilermakers"
  ) |>
  arrange(desc(team_gpg)) 
```

```{r}
ggplot() + geom_point(data=big_ten_gpg_2023, aes(x=total_games, y=total_goals)) +
  geom_point(data=umd_gpg_2023, aes(x=total_games, y=total_goals), color="red") +
  labs(
    title="In the Big Ten UMD scores low", 
    x="Total Goals", 
    y="Total Goals")
```

##Top 10

```{r}
ggplot() + 
  geom_bar(
    data=head(team_gpg_2023,10), 
    aes(x=reorder(team, team_gpg), weight=team_gpg)) + 
  coord_flip() + 
  scale_y_continuous(labels=comma) + 
  labs(
    title="Top Goal Per Game", 
    x="Team", 
    y="Goals per Game"
)
```

```{r}
ggplot() + 
  geom_bar(
    data=tail(team_gpg_2023,12), 
    aes(x=reorder(team, team_gpg), weight=team_gpg)) + 
  coord_flip() + 
  scale_y_continuous(labels=comma) + 
  labs(
    title="How many UMD Goals per Game?", 
    x="Team", 
    y="Goals per Game"
)
```

##What about the Freshman players?
```{r}
team_fr_goals_2023 <- player_goals_2023 |>
   filter(yr == "Fr") |>
   group_by(team)|>
  summarise(total_first_year_goals = sum(goals))|>
  arrange(desc(total_first_year_goals)) 
```

```{r}
umd_fr_goals_2023 <- team_fr_goals_2023 |> filter(team == "Maryland Terrapins, Terps")
```



```{r}
ggplot() + 
  geom_bar(
    data=head(team_fr_goals_2023, 10), 
    aes(x=reorder(team, total_first_year_goals), weight=total_first_year_goals)) + 
  coord_flip() + 
  scale_y_continuous(labels=comma) + 
  labs(
    title="Leading Freshman Scorers", 
    x="Team", 
    y="Goals per Game")
```


#Combining 
```{r}
combined <- team_gpg_2023 |> left_join(team_fr_goals_2023, join_by(team))
```


```{r}
team_fr_pct <- combined |>
 mutate(total_first_year_goals = replace_na(total_first_year_goals, 0)) |> 
  mutate(fr_pct = (total_first_year_goals/total_goals)*100)|>
  arrange(desc(total_first_year_goals))
```

```{r}
top_freshman_goals <- team_fr_pct |> filter(team =="Tex. A&M-Commerce Lions")
```

```{r}
two_top_freshman_goals <- team_fr_pct |> filter(team == "South Alabama Jaguars")
```


```{r}
umd_player_goals_2023 <- player_goals_2023 |> filter(team == "Maryland Terrapins, Terps")
```

```{r}
ggplot() + geom_point(data=team_fr_pct, aes(x=total_goals, y=total_first_year_goals)) +
  geom_smooth(data=team_fr_pct, aes(x=total_goals, y=total_first_year_goals), method="lm") + 
  scale_y_continuous(labels=comma) + 
  labs(
    title="UMD Does not re", 
    x="Total Goals", 
    y="Total First Year Goals") +
    geom_point(data=umd_gpg_2023, aes(x=total_games, y=total_goals), color="red") +
  geom_point(data=top_freshman_goals, aes(x=total_games, y=total_goals), color="green") +
  geom_point(data=two_top_freshman_goals, aes(x=total_games, y=total_goals), color="blue") 
```

Conclusion: No. Though there are teams that do such as Tex. A&M-Commerce Lions(green), and South Alabama Jaguars(blue). 












##What About Big Ten Teams?
```{r}
byyr <- ncaa_players_2023 |>
  group_by(team, yr) |>
  summarize(
    team_gpg = goals/games) |>
  filter(
    yr == "Fr" |
    yr == "Sr" |
    yr == "So" | 
    yr == "Jr") |>
  filter(
    team == "Ohio St. Buckeyes" |
    team == "Michigan St. Spartans" |
    team == "Penn St. Nittany Lions" |
    team == "Rutgers Scarlet Knights" |
    team == "Maryland Terrapins, Terps" |
    team == "Michigan St. Spartans" |
    team == "Indiana Hoosiers" |
    team == "Iowa Hawkeyes" |
    team == "Wisconsin Badgers" |
    team == "Minnesota Golden Gophers" |
    team == "Nebraska Huskers" |
    team == "Northwestern Wildcats" |
    team == "Illinois St. Redbirds" |
    team == "Purdue Boilermakers"
  )
```

Here I am creating a graph to demonstrate the relationship between each year for each of the Big 10 teams.The graph displays the amount of goals each of the years makes in comparison to one another on each team. 
```{r}
ggplot() + 
  geom_bar(
    data=byyr, 
    aes(x=reorder(team, team_gpg), weight=team_gpg, fill=yr)) + 
  coord_flip()
```


Here we can see how each year contributes to to goals per game for the season, in comparison with the each year and each team. 
Questions: Where in this data do you see a story? 
Are Freshman a significant contributor to a teams success?/Does this data need more support to answer that? 


